ActiVis: Mobile Object Detection and Active Guidance for People with Visual Impairments

### 2019-05-06 paper.tex
Merged param sections

ActiVis uses the camera's current and previous image data as input and leverages its understanding of inter-object spatial relationships to determine the best navigation action to take to reach the target object.
To enable this, we expanded upon our previous work and implemented a partially observable Markov Decision Process (POMDP)~\cite{bellman1957markovian} on a mobile phone that generates real-time navigation instructions to guide the user to their target object.
This work makes the following contributions:

@@

Section~\ref{sec:previous-work} discusses relevant previous work, followed by a description of the active vision system and controller in Section~\ref{sec:active-vision}.
The experiments that were conducted in this work and their results are discussed in Sections~\ref{sec:experiments} and~\ref{sec:results}, followed finally by a short conclusion and discussion of the next steps for the project in Section~\ref{sec:conclusion}.

@@

\subsection{Design}

@@

A POMDP model is represented by the 8-tuple $(\mathbf{S}, \mathbf{A}, \mathbf{T}, \mathbf{R}, \mathbf{\Omega}, \mathbf{O}, \mathbf{b}, \gamma)$, where $\mathbf{S}$ represents a finite set of discrete states, $\mathbf{A}$ is a set of discrete actions, $\mathbf{T}$ is a matrix containing the probabilities of transitioning from state $s$ to state $s'$, $s, s' \in \mathbf{S}$, after executing action $a$, $a \in \mathbf{A}$, and $\mathbf{R}$ is the reward the agent receives executing action $a$ and reaching state $s'$.
$\mathbf{\Omega}$ is the set of possible state observations, while the matrix $\mathbf{O}$ contains the probabilities of making state observation $\omega$, $\omega \in \mathbf{\Omega}$, when in state $s$ after executing action $a$.
Finally, $\mathbf{b}$ is the belief vector containing the state probability distribution and $\gamma$ is a discount factor that prioritises long-term over short-term rewards and affects the model's convergence rate. 
%$\mathbf{b}$ is updated for every timestep and action-observation pair.
Each of the POMDP's parameters are discussed next. 
%This update is done with

@@

\subsubsection{State Space}

The state is given by $s = \langle u, n, v \rangle$, where $u, u\in\mathbf{U}$, is the object within view, n the number of search steps taken since the search began and a binary variable $v$ that tracks the uniqueness of a waypoint's position for the search.

\subsubsection{Actions}

@@

\subsubsection{Transitions}

@@

\subsubsection{Reward Function}

@@

\begin{table}
  \centering
  \caption{The reward function for the POMDP. }\label{tab:rewards}
  \begin{tabular}{p{3cm}p{1cm}}
    \toprule
    $r(o = o_{target})$ & 10000 \\ \midrule
    $r(v = \textrm{true})$  & -75 \\ \midrule
    $r(n > n_{\max})$ & -75 \\ \midrule
    otherwise $r(\cdot)$ & -100  \\ \midrule
    \bottomrule
  \end{tabular}
\end{table}

\subsubsection{Observations and Observation Matrix}

@@

The POMDP model is put through a training process to generate a policy that contains the optimal belief-action mapping that the controller can use to produce the optimal waypoint locations.
This is done by having the model explore the entire state-action-observation space and optimising the policy to maximise its long-term cumulative reward.
The difficulty of solving POMDPs is that $\mathbf{b}$ is a vector of continuous probability distributions with infinite combinations, making POMDPs time-consuming or even impossible to solve exactly.
We found this to be the case for our moderately-sized state space size and therefore opted to use an approximate method instead.
Multiple algorithms have been proposed~\cite{bargiacchi2016dynamic,silver2010monte} and in this work we used the Point-Based Value Iteration (PBVI) algorithm~\cite{pineau2003point} as implemented in the AI-Toolbox library\footnote{github.com/Svalorzen/AI-Toolbox} to solve the POMDP model.
The PBVI algorithm speeds up the optimisation process by selecting a smaller subset of representative belief points from the belief function and tracking these points only. 
Using the PBVI algorithm, we generated a total of 15 policies, one for each target object.

@@

To describe the waypoints' pan and tilt positions, we implemented an improved version of the interface described in~\cite{bellotto2013} that uses a spatialised audio signal.
However, since our headphones bypass the ear's pinnae that allow a person to localise the height~\cite{roffler1968factors}, we spatialise the audio in the pan dimension only.
To convey the tilt angle, we instead exploit a human's natural association of high and low sound sources with a high and low sound frequency respectively~\cite{blauert1997spatial} and adjust the sound source's pitch accordingly. 
A similar approach was used in~\cite{schauerte2012assistive}.
With this interface and headset, ActiVis is able to describe location of a waypoint to a user in the pan and tilt dimensions.

@@

SSD-Lite is a single-stage object detection and classification network that is based on the SSD architecture that implements MobileNetV2~\cite{sandler2018mobilenetv2}.
This is a lightweight model that requires relatively little memory to perform inference tasks, making it suitable for mobile platforms. 
This model achieves a mAPs of 0.22 with 4.3M parameters and 0.8B FLOPS~\cite{li2018tinydsod} on the COCO dataset~\cite{lin2014microsoft}. 
The full SSD model achieves a slightly higher mAPs (0.25), but with significantly more parameters (34.3M) and FLOPS (34.36B).
%Tiny-DSOD~\cite{li2018tinydsod}, an alternative to SSD-Lite, has similar precision and computing requirements (mAPs=0.23, parameters=1.15M, FLOPS=1.12B), but we found that SSD-Lite produced better results for this application and chose to implement it in ActiVis.

>>>>

ActiVis uses the camera's current and previous image data as input and leverages its understanding of inter-object spatial relationships to determine the best navigation action to take to reach the target object.
To enable this, we expanded upon our previous work and implemented a partially observable Markov Decision Process (POMDP) on a mobile phone that generates real-time navigation instructions to guide the user to their target object.
This work makes the following contributions:

@@

Section~\ref{sec:previous-work} discusses relevant previous work, followed by a description of the active vision system and controller in Section~\ref{sec:active-vision}.
The experiments and their results are discussed in Sections~\ref{sec:experiments} and~\ref{sec:results}, followed finally by a short conclusion and discussion of the next steps for the project in Section~\ref{sec:conclusion}.

@@

\subsection{Controller Design}

@@

A POMDP model is represented by the 8-tuple $(\mathbf{S}, \mathbf{A}, \mathbf{T}, \mathbf{R}, \mathbf{\Omega}, \mathbf{O}, \mathbf{b}, \gamma)$, where $\mathbf{S}$ represents a finite set of discrete states, $\mathbf{A}$ is a set of discrete actions, $\mathbf{T}$ is a matrix containing the probabilities of transitioning from state $s$ to state $s'$, $s, s' \in \mathbf{S}$, after executing action $a$, $a \in \mathbf{A}$, and $\mathbf{R}$ is the reward the agent receives executing action $a$ and reaching state $s'$.
$\mathbf{\Omega}$ is the set of possible state observations, while the matrix $\mathbf{O}$ contains the probabilities of making state observation $\omega$, $\omega \in \mathbf{\Omega}$, when in state $s$ after executing action $a$.
Finally, $\mathbf{b}$ is the belief vector containing the state probability distribution and $\gamma$ is a discount factor that prioritises long-term over short-term rewards and affects the model's convergence rate. 
%$\mathbf{b}$ is updated for every timestep and action-observation pair.
The POMDP's parameters and training process are discussed next. 
%This update is done with

@@

\subsubsection{Parameters}

The state is given by $s = \langle u, n, v \rangle$, where $u, u\in\mathbf{U}$, is the object within view, $n$ the number of search steps taken during search and a binary variable $v$ that tracks the uniqueness of a waypoint's position for the search.

@@

@@

@@

\begin{table}[h]
  \centering
  \caption{The reward function for the POMDP. }\label{tab:rewards}
  \begin{tabular}{p{3cm}p{1cm}}
    \toprule
    $r(o = o_{target})$ & 10000 \\ \midrule
    $r(v = \textrm{true})$  & -75 \\ \midrule
    $r(n > n_{\max})$ & -75 \\ \midrule
    otherwise $r(\cdot)$ & -100  \\ \midrule
    \bottomrule
  \end{tabular}
\end{table}

@@

The POMDP model is put through a training process to generate a policy that contains the optimal belief-action mapping that the controller can use to produce the optimal waypoint locations.
This is done by having the model explore the entire state-action-observation space and optimising the policy to maximise its long-term cumulative reward.
The difficulty of solving POMDPs is that $\mathbf{b}$ is a vector of continuous probability distributions with infinite combinations, making POMDPs time-consuming or even impossible to solve exactly.
We found this to be the case for our moderately-sized state space size and therefore opted to use an approximate method instead.
In this work we used the Point-Based Value Iteration (PBVI) algorithm~\cite{pineau2003point} as implemented in the AI-Toolbox library\footnote{github.com/Svalorzen/AI-Toolbox} to solve the POMDP model.
The PBVI algorithm speeds up the optimisation process by selecting a smaller subset of representative belief points from the belief function and tracking these points only. 
Using the PBVI algorithm, we generated a total of 15 policies, one for each target object.

@@

To describe the waypoints' pan and tilt positions, we implemented an improved version of the interface described in~\cite{bellotto2013} that uses a spatialised audio signal.
However, since our headphones bypass the ear's pinnae that allow a person to localise the height~\cite{roffler1968factors}, we spatialise the audio in the pan dimension only.
To convey the tilt angle, we instead exploit a human's natural association of high and low sound sources with a high and low sound frequency respectively~\cite{blauert1997spatial} and adjust the sound source's pitch accordingly. 
%A similar approach was used in~\cite{schauerte2012assistive}.
With this interface and headset, ActiVis is able to describe location of a waypoint to a user in the pan and tilt dimensions.

@@

SSD-Lite is a single-stage object detection and classification network that is based on the SSD architecture that implements MobileNetV2~\cite{sandler2018mobilenetv2}.
This is a lightweight model that requires relatively little memory to perform inference tasks, making it suitable for mobile platforms. 
This model achieves a mAPs of 0.22 with 4.3M parameters and 0.8B FLOPS~\cite{li2018tinydsod} on the COCO dataset.%~\cite{lin2014microsoft}. 
The full SSD model achieves a slightly higher mAPs (0.25), but with significantly more parameters (34.3M) and FLOPS (34.36B).
%Tiny-DSOD~\cite{li2018tinydsod}, an alternative to SSD-Lite, has similar precision and computing requirements (mAPs=0.23, parameters=1.15M, FLOPS=1.12B), but we found that SSD-Lite produced better results for this application and chose to implement it in ActiVis.
